A Simple Minibatch Acceptance Test for MCMC

### 2016-05-17 nips_2016.tex
Adding a few references, modified Section 2. Now I will do Section 3, then work on experiments

Unfortunately, computing $f$ requires the use of all $N$ training data points. Moreover, it is
difficult to design tests using substantially fewer than $N$ points that also satisfy detailed
balance. There has been research on ways to conduct exact accept/reject tests using unbiased
estimators of the likelihood~\cite{Andrieu09thepseudo-marginal}. However, unbiased estimators of the
likelihood that can be computed from minibatches of data, such as the Kennedy-Bhanot
estimator~\cite{PhysRevD} have very high variance for large datasets. Because of this, once we get a
very high estimate of the likelihood, almost all proposed moves are rejected and the algorithm gets
stuck.

{\color{blue}
Daniel: I am worried. We need to motivate why it is not easy to use minibatch MCMC, and what other
researchers have tried to do about that.  The last paragraph is from~\cite{cutting_mh_2014}, and it
seems to argue that people have not used minibatch MCMC before because ``the algorithm gets stuck''.
But \emph{isn't that exactly what our algorithm is doing when it is getting ``stuck'' at modes}? If
we wanted to get stuck at modes with large datasets, why didn't we just do what previous work did?
The only way around this is if our algorithm is getting stuck at ``better'' modes, but then we need
to design our experiments to test this.
}

The work of~\cite{cutting_mh_2014} proposes an adaptive minibatch approach by using a sequential
hypothesis testing algorithm. During each iteration, they start with a small minibatch of data and
test the hypothesis that the sample $\theta'$ should be accepted or rejected based the likelihood
ratio. If the test cannot make a decision over a certain confidence threshold, then they increase
the minibatch size and test again. This process repeats until a decision.  The downside of this
algorithm is the need to keep incrementing the mini-batch size in one iteration; in the worst case,
all $N$ data points may be needed in a single iteration. A similar approach
from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties, but comes at the
cost of more computational time. The work of~\cite{conf/uai/MaclaurinA14} presents another minibatch
MCMC method, but which gets the \emph{exact} posterior, with the only requirement being that there
needs to be a lower bound on the log likelihood for each data point. They augment the likelihood
with auxiliary random variables, but this can hurt mixing.

@@

Minibatch SGHMC and Langevin Dynamics are orthogonal to our work; two ways of attempting
to solve the same problem of MCMC methods with big datasets. We could augment their methods with our
MH test, to get the benefit of their high quality proposals. Note that while their methods in theory
do not need an MH test, in practice, they need one due to discretization of the system. Furthermore,
    tests may also be needed for when the algorithms are run with a constant step size.

%% Daniel: I don't know how to integrate this easily.
%% \cite{sgmcmc_2015} (NIPS 2015) might want to give this a mention, where it would fit in their
%% recipe.

%% Daniel: for a further discussion on the 'Firefly' paper, see this blog post:
%% https://xianblog.wordpress.com/2014/04/02/firefly-monte-carlo/
%% Note: this guy also blogged about our earlier ICLR 2015 paper.

@@

\begin{figure}[ht]
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  %\includegraphics[width=0.5\linewidth]{empty}
  \caption{This will be a sequence of images for the Gaussian mixture.}
\end{figure}

@@

In this experiment, we use logistic regression to classify the MNIST dataset. For similification, we
only train the binary classifer for digits 1 vs 7. The dataset consisted of 12007 training data and
1000 testing data. Similarly, we use random walk as the proposer. We compare our mini-batch MH
testing with the adaptive mini-batch MH testing. Figure 4 shows the loglikelihood value and
perdiction accuracy over training process. Since the adaptive MH testing method needs to increase
the mini-batch size adaptively, we use the cumulative number of data instance each algorithm
consumes as the horizontal axis. We find our mini-batch MH testing can achieve the similar (or even
better loglikelihood/prediction accuracy) with consuming much fewer data points. Thus, we can assert
that our mini-batch MH testing is much more efficient than the adaptive one.

In this experiment, we also illustrate that by appling the temperature, we can achieve slight faster convergence than the constant temperature one.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/acc_hyc.png}
        \caption{prediction accuracy}
        \label{fig:logistic_pa}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/ll_hyc.png}
        \caption{loglikelihood}
        \label{fig:logistic_ll}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/roofline_hyc.png}
        \caption{roofline property}
        \label{fig:logistic_roof}
    \end{subfigure}
    \caption{Experiments of logistic regression}\label{fig:logistic}
\end{figure}

We use this experiment to illustarte another property of our mini-batch MH testing: when the size of mini-batch increase to a threshold, the accuracy / loglikelihood of sampled data at a fix iteration number will not improve with larger mini-batch size, see Fig.~4c. This is because the variance of $\Delta'$ decrease with large mini-batch, and the error of our mini-batch MH testing is proportional to our estimation error of variance of $\Delta'$. When the $var(\Delta')$ decrease to a threshold, the error of our mini-batch MH testing can be ignored.

{\color{blue}

For additional details, see Appendix~\ref{app:logistic}.
}

\subsection{Neural Network Optimization}\label{ssec:nets}
In this experiment, we apply about mini-batch MH testing with SGHMC proposer to sample the parameter instances from a fully connected neural network. We use Higgs data set from University of California Irvine. It is a classification problem to distinguish between a signal process which produces Higgs bosons and a background process which does not. We build a 4 layers neural network to tackle this problem. The first layer is a input layer, the last layer is a softmax layer, the intermediate layers are linear followed by Sigmoid, norm, dropout, starting and ending with linear operation.

We choose the 
\begin{figure}[ht]
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  %\includegraphics[width=0.5\linewidth]{empty}
  \caption{This will hopefully show results for neural networks on BIDMach.}
\end{figure}

@@

Straightforward directions for future work include more extensive experiments. More elaborate
extensions could be combining our results with Hamiltonian Monte Carlo methods, or integrating
parallel MCMC~\cite{conf/uai/AngelinoKWSA14} concepts into our work.

>>>>

Unfortunately, computing $f$ requires the use of all $N$ training data points. Moreover, it is
difficult to design tests using substantially fewer than $N$ points that also satisfy detailed
balance. To reconcile these competing objectives,~\cite{cutting_mh_2014} propose an adaptive
minibatch MCMC algorithm which uses a sequential hypothesis test. During each iteration, the
algorithm starts with a small minibatch of data and tests the hypothesis that the sample $\theta'$
should be accepted or rejected based on the likelihood ratio. If the test cannot make a decision
over a certain confidence threshold, then they increase the minibatch size and test again. This
process repeats until a decision.  The downside of this algorithm is the need to keep incrementing
the mini-batch size in one iteration; in the worst case, all $N$ data points may be needed in a
single iteration.

A similar approach from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical
properties, but comes at the cost of more computational time. The work
of~\cite{conf/uai/MaclaurinA14} presents another minibatch MCMC method, but which gets the
\emph{exact} posterior, with the only requirement being that there needs to be a lower bound on the
log likelihood for each data point. They augment the likelihood with auxiliary random variables, but
this can hurt mixing.

@@

Minibatch SGHMC and Langevin Dynamics are orthogonal to our work; two ways of attempting
to solve the same problem of MCMC methods with big datasets. We could augment their methods with our
MH test, to get the benefit of their high quality proposals. Note that while their methods in theory
do not need an MH test, in practice, they need one due to discretization of the system. Furthermore,
tests may also be needed for when the algorithms are run with a constant step size. {\color{blue}
Daniel: not totally confident on this paragraph, how much SGHMC/LD do we discuss?}

%% Daniel: for a further discussion on the 'Firefly' paper, see this blog post:
%% https://xianblog.wordpress.com/2014/04/02/firefly-monte-carlo/
%% Note: this guy also blogged about our earlier ICLR 2015 paper. He's the third author of the SAME
%% paper, which explains why he knows this stuff.

@@

\begin{figure}[t]
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  %\includegraphics[width=0.5\linewidth]{empty}
  \caption{This will be a sequence of images for the Gaussian mixture.}
\end{figure}

@@

In this experiment, we use logistic regression to classify the MNIST
dataset~\cite{lecun-mnisthandwrittendigit-2010}. For simplification, we only train the binary
classifier for digits 1 vs 7. The dataset consisted of 12007 training data and 1000 testing data.
Similarly, we use random walk as the proposer. We compare our mini-batch MH testing with the
adaptive mini-batch MH testing. Figure 4 shows the log likelihood value and prediction accuracy over
training process. Since the adaptive MH testing method needs to increase the mini-batch size
adaptively, we use the cumulative number of data instance each algorithm consumes as the horizontal
axis. We find our mini-batch MH testing can achieve the similar (or even better log
likelihood/prediction accuracy) with consuming much fewer data points. Thus, we can assert that our
mini-batch MH testing is much more efficient than the adaptive one.

In this experiment, we also illustrate that by applying the temperature, we can achieve slight
faster convergence than the constant temperature one.

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/acc_hyc.png}
        \caption{prediction accuracy}
        \label{fig:logistic_pa}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/ll_hyc.png}
        \caption{loglikelihood}
        \label{fig:logistic_ll}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{./figures/roofline_hyc.png}
        \caption{roofline property}
        \label{fig:logistic_roof}
    \end{subfigure}
    \caption{Experiments of logistic regression}\label{fig:logistic}
\end{figure}

We use this experiment to illustrate another property of our mini-batch MH testing: when the size of
mini-batch increase to a threshold, the accuracy / log likelihood of sampled data at a fix iteration
number will not improve with larger mini-batch size, see Fig.~4c. This is because the variance of
$\Delta'$ decrease with large mini-batch, and the error of our mini-batch MH testing is proportional
to our estimation error of variance of $\Delta'$. When the $var(\Delta')$ decrease to a threshold,
the error of our mini-batch MH testing can be ignored.

{\color{blue}
Daniel/Haoyu: For additional details, see Appendix~\ref{app:logistic}.
}

\subsection{Neural Network Optimization}\label{ssec:nets}

In this experiment, we apply about mini-batch MH testing with SGHMC proposer to sample the parameter instances from a fully connected neural network. We use Higgs data set from University of California Irvine. It is a classification problem to distinguish between a signal process which produces Higgs bosons and a background process which does not. We build a 4 layers neural network to tackle this problem. The first layer is a input layer, the last layer is a softmax layer, the intermediate layers are linear followed by Sigmoid, norm, dropout, starting and ending with linear operation.

We choose the 
\begin{figure}[t]
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  %\includegraphics[width=0.5\linewidth]{empty}
  \caption{This will hopefully show results for neural networks on BIDMach.}
\end{figure}

@@

In this paper, we have derived a new MH test for minibatch MCMC methods. We demonstrated how a
simple deconvolution process allows us to use a minibatch approximation to the full data tests. We
experimentally show the benefits our MH test.

Straightforward directions for future work include running additional experiments, with a particular
focus on investigation of the variance precondition.  More elaborate extensions include combining
our results with Hamiltonian Monte Carlo methods, providing a recipe for how to use our algorithm
(following the general framework of~\cite{sgmcmc_2015}), or integrating parallel
MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts into our work.

{\color{blue}
Daniel: TODO, clean this up.
}
